rr

### 2025-11-07 double_elimination.qmd

```{r}
simulate_double_elimination_tournament <- function(num_teams,distribution,seeding_structure = NULL, ties=T, series=1) {

  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }
  teams <- paste(num_teams:1)
  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }
  if (distribution == "Normal"){
    strengths <- sapply(num_teams, function(n) { qnorm(1:n/(n+1)) })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  }
  else if (distribution == "Uniform"){
    strengths <- sapply(num_teams, function(n) { qunif(1:n/(n+1) , 0 , sqrt(12)) })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams)),
      rank_hat = rep(NA,length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  }
     else if (distribution == "Manual") {
    cat("Please enter the strengths for each team in descending rank order:\n")
    strengths <- numeric(num_teams)
    for (i in num_teams:1) {
      repeat {
        input <- as.numeric(readline(paste("Strength for Team", num_teams - i + 1, ": ")))
        if (!is.na(input)) {
          strengths[i] <- input
          break
        } else {
          cat("Invalid input. Please enter a numeric value.\n")
        }
      }
    }
    manual_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(manual_strengths, true_rank)
  }
  else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)){
    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
  } else{
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }
  if(series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

@@

    team_rank <- num_teams
    winners_bracket <- teams
    strengths <- permuted_df$true_strength

    losers_bracket <- c()

    # gets results of the winners bracket
    while (length(winners_bracket) > 1 || length(losers_bracket > 0)) {
      eliminated <- c()
      if (length(winners_bracket) > 1) {
      #cat("\n--- Winner's Bracket --- \n")
      next_winners <- c() 
      round_losers <- c()

      for (j in seq(1, length(winners_bracket), by = 2)) { # Gets 2 teams for each match
        if (j + 1 > length(winners_bracket)) break
        team1 <- winners_bracket[j]
        team2 <- winners_bracket[j+1]

        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

        wins_team1 <- 0
        wins_team2 <- 0
        losses_team1 <- 0
        losses_team2 <- 0
        for (game in 1:series){
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          cat(sprintf("Matchup:  %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))
          if(!is.na(team1)) {
          if (match_winner == team1) {
            wins_team1 <- wins_team1 + 1
            losses_team2 <- losses_team2 + 1
          } else {
            wins_team2 <- wins_team2 + 1
            losses_team1 <- losses_team1 + 1
          }
          }
          else {wins_team2 <- (series+1) / 2}

          # Determine the overall winner if a majority is reached
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }
        # Update the game_wins and game_losses outside of the loop
        if (!any(is.na(c(team1, team2)))) {
          cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
        permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2,
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
        }
        # Determine the match winner based on series results
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
        series_loser <- ifelse(series_winner == team1, team2, team1)
        cat(sprintf("Matchup:  %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))
        next_winners <- c(next_winners, series_winner)
        round_losers <- c(round_losers, series_loser)
    }
     winners_bracket <- next_winners  # Winners move to the next round
     losers_bracket <- c(losers_bracket, round_losers)
     if(length(losers_bracket) %% 2 != 0){losers_bracket <- c(losers_bracket, NA)}

      }

    # Losers' bracket
    if (length(losers_bracket) > 0) {
      if(length(losers_bracket) > 1){
      cat("\n--- Loser's Bracket --- \n")
      }
      next_losers <- c()

      # alternates games for losers bracket
      if(length(losers_bracket) > 2){
        if(length(losers_bracket) == 4 && is.na(losers_bracket[4])){
          losers_bracket <- losers_bracket
        } else{
          top_bracket <- losers_bracket[seq(1,length(losers_bracket), by = 2)]
          bottom_bracket <- losers_bracket[seq(2,length(losers_bracket), by = 2)]
          losers_bracket <- c(top_bracket,bottom_bracket)
        }
      }

      for (j in seq(1, length(losers_bracket), by = 2)) {
        if (j + 1 > length(losers_bracket)) break
        team1 <- losers_bracket[j]
        team2 <- losers_bracket[j + 1]
        strength1 <- strengths[which(permuted_df$true_rank == team1)]
        strength2 <- strengths[which(permuted_df$true_rank == team2)]

@@

        # Play series
        for (game in 1:series) {
          match_winner <- simulate_match(team1, team2, strength1, strength2)
          if(is.na(match_winner)){
            wins_team1 == wins_team1 + 1
          }
          if(!is.na(team1)) {
            if (match_winner == team1) {
              wins_team1 <- wins_team1 + 1
              losses_team2 <- losses_team2 + 1
            } else {
              wins_team2 <- wins_team2 + 1
              losses_team1 <- losses_team1 + 1
            }
          }
          else {wins_team2 <- (series+1) / 2}

          # Determine overall series winner
          if (wins_team1 > series / 2 || wins_team2 > series / 2) {
            break
          }
        }
        if (!any(is.na(c(team1, team2)))) {
          cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
          permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2,
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
        }

        if(is.na(team1) && is.na(team2)){
          series_winner <- NA
          next_losers <- c(next_losers, NA)
        } else{
        series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
        series_loser <- ifelse(wins_team1 > wins_team2, team2, team1)
        next_losers <- c(next_losers, series_winner)
        eliminated <- c(eliminated, series_loser)
        cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, series_winner))
        }
      }
        team_rank <- team_rank - 1
        permuted_df$rank_hat[permuted_df$true_rank %in% eliminated] <- team_rank
        losers_bracket <- next_losers

    }

      # Final match for loser's bracket (aka third place game)
if (length(winners_bracket) == 1 && length(losers_bracket) == 1) {

# Final matchup between winners and losers
    cat("\n--- Final Match ---\n")
    final_match <- c(winners_bracket, losers_bracket)
    if(length(final_match) == 2){
    team1 <- final_match[1]
    team2 <- final_match[2]
    strength1 <- strengths[which(permuted_df$true_rank == team1)]
    strength2 <- strengths[which(permuted_df$true_rank == team2)]

@@

    for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
        losses_team2 <- losses_team2 + 1
      } else {
        wins_team2 <- wins_team2 + 1
        losses_team1 <- losses_team1 + 1
      }

      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
        break
      }
    }
    if (!any(is.na(c(team1, team2)))) {
      cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
      permuted_df <- permuted_df %>%
        mutate(game_wins = case_when(
          true_rank == team1 ~ game_wins + wins_team1,
          true_rank == team2 ~ game_wins + wins_team2,
          TRUE ~ game_wins
        )) %>%
        mutate(game_losses = case_when(
          true_rank == team1 ~ game_losses + losses_team1,
          true_rank == team2 ~ game_losses + losses_team2,
          TRUE ~ game_losses
          ))
    }

    if (wins_team1 > wins_team2) {
      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
    } else {
      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
      # second game if the loser's bracket team wins
      for (game in 1:series) {
      match_winner <- simulate_match(team1, team2, strength1, strength2)
      if (match_winner == team1) {
        wins_team1 <- wins_team1 + 1
        losses_team2 <- losses_team2 + 1
      } else {
        wins_team2 <- wins_team2 + 1
        losses_team1 <- losses_team1 + 1
      }

      if (wins_team1 > series / 2 || wins_team2 > series / 2) {
        break
      }
    }
      if (!any(is.na(c(team1, team2)))) {
        cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
        permuted_df <- permuted_df %>%
          mutate(game_wins = case_when(
            true_rank == team1 ~ game_wins + wins_team1,
            true_rank == team2 ~ game_wins + wins_team2 - 1, # wins_team2 = 2 in loop but only need to add one game
            TRUE ~ game_wins
          )) %>%
          mutate(game_losses = case_when(
            true_rank == team1 ~ game_losses + losses_team1 - 1,
            true_rank == team2 ~ game_losses + losses_team2,
            TRUE ~ game_losses
          ))
      }
      if (wins_team1 > wins_team2) {
      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team1))
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
    } else {
      cat(sprintf("Matchup:  %s vs %s | Winner: %s\n:", team1, team2, team2))
      permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
      permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
        }
      }
    }
  }
}
    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
    results[[i]] <- permuted_df
}
  final_results <- do.call(rbind, results)

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(true_rank) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
  }
  else{
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

  final_results <- final_results %>%
    mutate(distribution = distribution) %>%
    mutate(true_rank = as.numeric(true_rank)) %>%
    arrange(true_rank)
  return(final_results)
}

test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1)
#test
#manual_test <- simulate_double_elimination_tournament(num_teams = 4, distribution = "Manual", ties=F, series=1)
#manual_test

>>>>

```{r}
simulate_double_elimination_tournament <- function(num_teams, distribution, seeding_structure = NULL, ties = T, series = 1, final = T, theta_hat = NULL) {
  if (num_teams <= 3) {
    stop("Number of Teams must be greater than 3.")
  }

  teams <- paste(num_teams:1)

  if (log2(num_teams) %% 1 != 0) {
    power_of_2 <- 2^ceiling(log2(num_teams))
    teams <- c(teams, rep(NA, power_of_2 - num_teams))
  }

  if (distribution == "Normal") {
    strengths <- sapply(num_teams, function(n) {
      qnorm(1:n/(n+1))
    })
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  } else if (distribution == "Uniform") {
    strengths <- sapply(num_teams, function(n) {
      qunif(1:n/(n+1), 0, sqrt(12))
    })
    unif_strength <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams)),
      rank_hat = rep(NA, length(teams))
    )
    df <- arrange(unif_strength, true_rank)
  } else if (distribution == "Same"){
    strengths <- sapply(num_teams, function(n) {rep(0,n)})
    normal_strengths <- data.frame(
      true_rank = as.numeric(teams),
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA,length(teams)),
      game_wins = rep(0,length(teams)),
      game_losses = rep(0,length(teams))
    )
    df <- arrange(normal_strengths, true_rank)
  } else if (distribution == "Manual") {
    strengths <- theta_hat
    manual_strengths <- data.frame(
      true_strength = c(strengths, rep(NA, length(teams) - num_teams)),
      rank_hat = rep(NA, length(teams)),
      game_wins = rep(0, length(teams)),
      game_losses = rep(0, length(teams))
    )
    df <- manual_strengths %>%
      arrange(true_strength) %>%
      mutate(true_rank = as.numeric(teams)) %>%
      arrange(true_strength, true_rank)
  } else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

  # Sets seed to that of common best vs worst structure
  if (is.null(seeding_structure)) {
    seeding_structure <- matrix(generate_elimination_tournament_structure(num_teams), nrow = 1)
  } else {
    seeding_structure <- matrix(seeding_structure, nrow = 1)
  }

  if (series %% 2 == 0) {
    stop("Series must be an odd number to avoid potential ties.")
  }

@@

    team_rank <- num_teams
    winners_bracket <- teams
    strengths <- permuted_df$true_strength
    losers_bracket <- c()

    # gets results of the winners bracket
    while (length(winners_bracket) > 1 || length(losers_bracket > 0)) {
      eliminated <- c()

      if (length(winners_bracket) > 1) {
        #cat("\n--- Winner's Bracket --- \n")
        next_winners <- c()
        round_losers <- c()

        for (j in seq(1, length(winners_bracket), by = 2)) {
          # Gets 2 teams for each match
          if (j + 1 > length(winners_bracket)) break

          team1 <- winners_bracket[j]
          team2 <- winners_bracket[j+1]
          strength1 <- strengths[which(permuted_df$true_rank == team1)]
          strength2 <- strengths[which(permuted_df$true_rank == team2)]

@@

          for (game in 1:series) {
            match_winner <- simulate_match(team1, team2, strength1, strength2)
            #cat(sprintf("Matchup: %s vs %s | Strengths: %.2f vs %.2f | Winner: %s\n", team1, team2, strength1, strength2, match_winner))

            if (!is.na(team1)) {
              if (match_winner == team1) {
                wins_team1 <- wins_team1 + 1
                losses_team2 <- losses_team2 + 1
              } else {
                wins_team2 <- wins_team2 + 1
                losses_team1 <- losses_team1 + 1
              }
            } else {
              wins_team2 <- (series+1) / 2
            }

            # Determine the overall winner if a majority is reached
            if (wins_team1 > series / 2 || wins_team2 > series / 2) {
              break
            }
          }

          # Update the game_wins and game_losses outside of the loop
          if (!any(is.na(c(team1, team2)))) {
            #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
            permuted_df <- permuted_df %>%
              mutate(game_wins = case_when(
                true_rank == team1 ~ game_wins + wins_team1,
                true_rank == team2 ~ game_wins + wins_team2,
                TRUE ~ game_wins
              )) %>%
              mutate(game_losses = case_when(
                true_rank == team1 ~ game_losses + losses_team1,
                true_rank == team2 ~ game_losses + losses_team2,
                TRUE ~ game_losses
              ))
          }

          # Determine the match winner based on series results
          series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
          series_loser <- ifelse(series_winner == team1, team2, team1)
          #cat(sprintf("Matchup: %s vs %s | Winner: %s\n: | Loser: %s\n", team1, team2, series_winner, series_loser))

          next_winners <- c(next_winners, series_winner)
          round_losers <- c(round_losers, series_loser)
        }

        winners_bracket <- next_winners # Winners move to the next round
        losers_bracket <- c(losers_bracket, round_losers)
        if (length(losers_bracket) %% 2 != 0) {
          losers_bracket <- c(losers_bracket, NA)
        }
      }

      # Losers' bracket
      if (length(losers_bracket) > 0) {
        if (length(losers_bracket) > 1) {
          #cat("\n--- Loser's Bracket --- \n")
        }
        next_losers <- c()

        # alternates games for losers bracket
        if (length(losers_bracket) > 2) {
          if (length(losers_bracket) == 4 && is.na(losers_bracket[4])) {
            losers_bracket <- losers_bracket
          } else {
            top_bracket <- losers_bracket[seq(1, length(losers_bracket), by = 2)]
            bottom_bracket <- losers_bracket[seq(2, length(losers_bracket), by = 2)]
            losers_bracket <- c(top_bracket, bottom_bracket)
          }
        }

        for (j in seq(1, length(losers_bracket), by = 2)) {
          if (j + 1 > length(losers_bracket)) break

          team1 <- losers_bracket[j]
          team2 <- losers_bracket[j + 1]
          strength1 <- strengths[which(permuted_df$true_rank == team1)]
          strength2 <- strengths[which(permuted_df$true_rank == team2)]

@@

          # Play series
          for (game in 1:series) {
            match_winner <- simulate_match(team1, team2, strength1, strength2)
            if (is.na(match_winner)) {
              wins_team1 == wins_team1 + 1
            }

            if (!is.na(team1)) {
              if (match_winner == team1) {
                wins_team1 <- wins_team1 + 1
                losses_team2 <- losses_team2 + 1
              } else {
                wins_team2 <- wins_team2 + 1
                losses_team1 <- losses_team1 + 1
              }
            } else {
              wins_team2 <- (series+1) / 2
            }

            # Determine overall series winner
            if (wins_team1 > series / 2 || wins_team2 > series / 2) {
              break
            }
          }

          if (!any(is.na(c(team1, team2)))) {
            #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
            permuted_df <- permuted_df %>%
              mutate(game_wins = case_when(
                true_rank == team1 ~ game_wins + wins_team1,
                true_rank == team2 ~ game_wins + wins_team2,
                TRUE ~ game_wins
              )) %>%
              mutate(game_losses = case_when(
                true_rank == team1 ~ game_losses + losses_team1,
                true_rank == team2 ~ game_losses + losses_team2,
                TRUE ~ game_losses
              ))
          }

          if (is.na(team1) && is.na(team2)) {
            series_winner <- NA
            next_losers <- c(next_losers, NA)
          } else {
            series_winner <- ifelse(wins_team1 > wins_team2, team1, team2)
            series_loser <- ifelse(wins_team1 > wins_team2, team2, team1)
            next_losers <- c(next_losers, series_winner)
            eliminated <- c(eliminated, series_loser)
            #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, series_winner))
          }
        }

        team_rank <- team_rank - 1
        permuted_df$rank_hat[permuted_df$true_rank %in% eliminated] <- team_rank
        losers_bracket <- next_losers
      }

      # Final match for loser's bracket (aka third place game)
      if (length(winners_bracket) == 1 && length(losers_bracket) == 1) {
        if (final == F){
          # No final game played between winner of winner bracket and winner of loser bracket
          team1 <- winners_bracket[1]
          team2 <- losers_bracket[1]
          permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
          permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
          break
        }
        # Final matchup between winners and losers
        else if (final == T) {
          #cat("\n--- Final Match ---\n")
          final_match <- c(winners_bracket, losers_bracket)

          if (length(final_match) == 2) {
            team1 <- final_match[1]
            team2 <- final_match[2]
            strength1 <- strengths[which(permuted_df$true_rank == team1)]
            strength2 <- strengths[which(permuted_df$true_rank == team2)]

            for (game in 1:series) {
              match_winner <- simulate_match(team1, team2, strength1, strength2)
              if (match_winner == team1) {
                wins_team1 <- wins_team1 + 1
                losses_team2 <- losses_team2 + 1
              } else {
                wins_team2 <- wins_team2 + 1
                losses_team1 <- losses_team1 + 1
              }
              if (wins_team1 > series / 2 || wins_team2 > series / 2) {
                break
              }
            }

            if (!any(is.na(c(team1, team2)))) {
              #cat(sprintf("Losses - %s: %d | %s: %d\n", team1, losses_team1, team2, losses_team2))
              permuted_df <- permuted_df %>%
                mutate(game_wins = case_when(
                  true_rank == team1 ~ game_wins + wins_team1,
                  true_rank == team2 ~ game_wins + wins_team2,
                  TRUE ~ game_wins
                )) %>%
                mutate(game_losses = case_when(
                  true_rank == team1 ~ game_losses + losses_team1,
                  true_rank == team2 ~ game_losses + losses_team2,
                  TRUE ~ game_losses
                ))
            }

            if (wins_team1 > wins_team2) {
              #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team1))
              permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
              permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
            } else {
              #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team2))
              # second game if the loser's bracket team wins
              # RESET the counters for the second series
              wins_team1 <- 0
              wins_team2 <- 0
              losses_team1 <- 0
              losses_team2 <- 0

              for (game in 1:series) {
                match_winner <- simulate_match(team1, team2, strength1, strength2)
                if (match_winner == team1) {
                  wins_team1 <- wins_team1 + 1
                  losses_team2 <- losses_team2 + 1
                } else {
                  wins_team2 <- wins_team2 + 1
                  losses_team1 <- losses_team1 + 1
                }
                if (wins_team1 > series / 2 || wins_team2 > series / 2) {
                  break
                }
              }

              if (!any(is.na(c(team1, team2)))) {
                #cat(sprintf("Wins - %s: %d | %s: %d\n", team1, wins_team1, team2, wins_team2))
                permuted_df <- permuted_df %>%
                  mutate(game_wins = case_when(
                    true_rank == team1 ~ game_wins + wins_team1,
                    true_rank == team2 ~ game_wins + wins_team2,
                    TRUE ~ game_wins
                  )) %>%
                  mutate(game_losses = case_when(
                    true_rank == team1 ~ game_losses + losses_team1,
                    true_rank == team2 ~ game_losses + losses_team2,
                    TRUE ~ game_losses
                  ))
              }

              if (wins_team1 > wins_team2) {
                #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team1))
                permuted_df$rank_hat[permuted_df$true_rank == team1] <- 1
                permuted_df$rank_hat[permuted_df$true_rank == team2] <- 2
              } else {
                #cat(sprintf("Matchup: %s vs %s | Winner: %s\n:", team1, team2, team2))
                permuted_df$rank_hat[permuted_df$true_rank == team2] <- 1
                permuted_df$rank_hat[permuted_df$true_rank == team1] <- 2
              }
            }
          }
        }
      }
    }

    permuted_df$seed <- rep(paste(seeding_structure[i, ], collapse = ", "), nrow(permuted_df))
    results[[i]] <- permuted_df
  }

  final_results <- do.call(rbind, results)

  # Remove ties if ties = FALSE
  if (!ties) {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(true_rank) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "random"))
  } else {
    final_results <- final_results %>%
      filter(!is.na(true_rank)) %>%
      arrange(rank_hat) %>%
      mutate(rank_hat = rank(rank_hat, ties.method = "average"))
  }

  final_results <- final_results %>%
    mutate(distribution = distribution) %>%
    mutate(true_rank = as.numeric(true_rank)) %>%
    arrange(true_rank)

  return(final_results)
}

test <- simulate_double_elimination_tournament(num_teams = 8, distribution = "Normal", ties=F, series=1, final = F)
#test
#manual_test <- simulate_double_elimination_tournament(num_teams = 4, distribution = "Manual", ties=F, series=1)
#manual_test

### 2025-11-07 round_robin.qmd

```{r}
simulate_roundrobin_r <- function(teams, strengths, r){
  n <- length(teams)
  win_counts <- numeric(n)
  for (i in 1:r){
    for (i in seq_along(teams)){
      for (j in seq_along(teams)){
        if (i != j){
          match_winner <- simulate_match(teams[i],teams[j],strengths)
          if (match_winner == teams[i]) {
            win_counts[i] <- win_counts[i] + 1
          } else {
            win_counts[j] <- win_counts[j] + 1
          }
        }
      }
    }
  }
  results <- data.frame(
    Team = teams,
    Game_Wins = win_counts
  )

  # Ranks teams based on number of wins (includes ties)
  results$Rank_Hat <- rank(-results$Game_Wins, ties.method = "average")
  results <- results[order(results$Rank_Hat),]

  results <- results %>% 
    inner_join(strengths, by = 'Team') %>% 
    select(Team, True_Rank, True_Strength, Rank_Hat, Game_Wins)

>>>>

```{r}
simulate_roundrobin_r <- function(num_teams, distribution, ties = TRUE, r = 1, theta_hat = NULL) {
  teams <- paste(1:num_teams)
  n <- length(teams)

  # Initialize strength vector
  if (distribution == "Normal") {
    strengths <- qnorm(1:num_teams / (num_teams + 1))
  } else if (distribution == "Same") {
    strengths <- rep(0, num_teams)
  } else if (distribution == "Uniform") {
    strengths <- qunif(1:num_teams / (num_teams + 1), 0, sqrt(12))
  } else if (distribution == "Manual") {
    if (is.null(theta_hat)) {
      stop("theta_hat must be provided for Manual distribution")
    }
    strengths <- theta_hat
  } else {
    stop("Distribution not found: Enter Manual to input your own strengths")
  }

  # Initialize win/loss counters (indexed by original team position)
  game_wins <- rep(0, num_teams)
  game_losses <- rep(0, num_teams)

  # Simulate round robin tournament r times
  for (round in 1:r) {
    for (i in 1:(num_teams - 1)) {
      for (j in (i + 1):num_teams) {
        team1 <- teams[i]
        team2 <- teams[j]
        strength1 <- strengths[i]
        strength2 <- strengths[j]

        # Simulate match between team i and team j using their strengths
        match_winner <- simulate_match(team1, team2, strength1, strength2)

        # Check which team won based on the returned team name
        if (match_winner == team1) {
          game_wins[i] <- game_wins[i] + 1
          game_losses[j] <- game_losses[j] + 1
        } else {
          game_wins[j] <- game_wins[j] + 1
          game_losses[i] <- game_losses[i] + 1
        }
      }
    }
  }

  # Create final dataframe with results
  df <- data.frame(
    team = teams,
    true_strength = strengths,
    game_wins = game_wins,
    game_losses = game_losses,
    distribution = distribution
  )

  # Sort by true_strength to assign true rankings
  df <- df %>% 
    arrange(desc(true_strength)) %>% 
    mutate(true_rank = row_number())

  # Rank teams based on number of wins
  if (ties == TRUE) {
    df$rank_hat <- rank(-df$game_wins, ties.method = "average")
  } else {
    df$rank_hat <- rank(-df$game_wins, ties.method = "random")
  }

  # Select and order columns as specified
  df <- df %>% 
    select(true_rank, true_strength, rank_hat, game_wins, game_losses, distribution) %>%
    arrange(true_rank)

  return(df)
}

simulate_roundrobin_r(16,"Normal", r = 4)
```
